---
title: "Project"
author: "Manvir Singh, Saurav Mawandia, Sougandh Kohli"
date: "`r Sys.Date()`"
output: html_document
---

# Tables and Figures
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(readxl)
library(plyr)
library(ggplot2)


dataset = read.csv("~/Documents/545/545/dataset.xls")


summary(dataset)
# sec 2
data_set <- rename(dataset, c("sales" = "role"))

data_set$DateofHire <- as.Date(data_set$DateofHire, '%m/%d/%Y')
data_set$employementEndDateOrTodaysDate <- as.Date(
  ifelse(
    data_set$DateofTermination == '',
    '11/18/2021',
    data_set$DateofTermination
  ),
  '%m/%d/%Y'
)
data_set$exp_in_company <-
  round(
    difftime(
      data_set$employementEndDateOrTodaysDate,
      data_set$DateofHire,
      units = "days"
    ) / 365,
    digits = 3
  )

data_set$exp_in_company <- as.numeric(data_set$exp_in_company)
names(data_set)[10] <- "salary"
head(data_set)
dim(data_set)
str(data_set)


attrition <- as.factor(data_set$Termd)
summary(attrition)

perc_attrition_rate <- sum(data_set$Termd / length(data_set$Termd)) * 100
#percentage of attrition
print(perc_attrition_rate)

data_set$PerformanceScoreNumeric <-
  ifelse(data_set$PerformanceScore == 'Exceeds', 4, 0)

data_set$PerformanceScoreNumeric <-
  ifelse(data_set$PerformanceScore == 'Fully Meets',
         3,
         data_set$PerformanceScoreNumeric)
data_set$PerformanceScoreNumeric <-
  ifelse(
    data_set$PerformanceScore == 'Needs Improvement',
    2,
    data_set$PerformanceScoreNumeric
  )
data_set$PerformanceScoreNumeric <-
  ifelse(data_set$PerformanceScore == 'PIP',
         1,
         data_set$PerformanceScoreNumeric)



# Overview of summary (Turnover V.S. Non-turnover)

cor_vars <-
  data_set[, c(
    'SpecialProjectsCount',
    'PerformanceScoreNumeric',
    'EmpSatisfaction',
    'DaysLateLast30',
    'exp_in_company',
    'EngagementSurvey',
    'Termd'
  )]


aggregate(cor_vars[, c(
  'SpecialProjectsCount',
  'PerformanceScoreNumeric',
  'EmpSatisfaction',
  'DaysLateLast30',
  'exp_in_company',
  'EngagementSurvey'
)],
by = list(Category = cor_vars$Termd),
FUN = mean)
```
3b. Correlation Matrix & Heatmap¶

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(reshape2)
library(ggplot2)
cor_vars <-
  data_set[, c(
    'SpecialProjectsCount',
    'PerformanceScoreNumeric',
    'EmpSatisfaction',
    'DaysLateLast30',
    'exp_in_company',
    'EngagementSurvey',
    'Termd'
  )]
cor(cor_vars)
trans<-cor(cor_vars)
melted_cormat <- melt(trans)

ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

3b2. Statistical Test for Correlation
```{r echo=FALSE, message=FALSE, warning=FALSE}

# Let's compare the means of our employee turnover satisfaction against the employee population satisfaction
emp_population_satisfaction <-mean(data_set$EmpSatisfaction)
left_pop<-subset(data_set,Termd==1)

emp_turnover_satisfaction <-mean(left_pop$EmpSatisfaction)

print( c('The mean for the employee population is: ', emp_population_satisfaction) )
print( c('The mean for the employees that had a turnover is: ' ,emp_turnover_satisfaction) )
```
Conducting the T-Test

Let's conduct a t-test at 95% confidence level and see if it correctly rejects the null hypothesis that the sample comes from the same distribution as the employee population. To conduct a one sample t-test, we can use the stats.ttest_1samp() function:

```{r echo=FALSE, message=FALSE, warning=FALSE}
t.test(left_pop$EmpSatisfaction,mu=emp_population_satisfaction)
```



T-Test Quantile
```{r echo=FALSE, message=FALSE, warning=FALSE}
#degress of freedom
dof<-sum(as.numeric(data_set$Termd))

LQ <-qt(0.025,dof)  # Left Quartile

RQ <-qt(0.975,dof)  # Right Quartile

print (c('The t-distribution left quartile range is: ',LQ))
print (c('The t-distribution right quartile range is: ' ,RQ))
```

3c. Distribution Plots (Performance Score - Evaluation - AverageMonthlyHours)


```{r echo=FALSE, message=FALSE, warning=FALSE}
vis_1<-table(data_set$PerformanceScore,data_set$Termd)
#print(vis_1)
d_vis_1<-as.data.frame(vis_1)
print(d_vis_1)
library(ggplot2)
p<-ggplot(d_vis_1, aes(x=Var1,y=Freq,fill=Var2)) +
 geom_bar(position="dodge",stat='identity') + coord_flip()

print(p)
```
Position and Termination 
```{r echo=FALSE, message=FALSE, warning=FALSE}
vis_2<-table(data_set$Position,data_set$Termd)
d_vis_2<-as.data.frame(vis_2)
d_vis_2<-subset(d_vis_2,Var2==1)
#print(d_vis_2)
d_vis_2$Var1 <- factor(d_vis_2$Var1, levels = d_vis_2$Var1[order(-d_vis_2$Freq)])
p<-ggplot(d_vis_2, aes(x=Var1,y=Freq,fill=Var1)) +
 geom_bar(stat='identity') +theme(axis.text.x = element_text(angle = 90, hjust = 1))

print(p)
```


Race and Termination

```{r echo=FALSE, message=FALSE, warning=FALSE}
vis_3<-table(data_set$RaceDesc,data_set$Termd)
d_vis_3<-as.data.frame(vis_3)
#print(d_vis_1)
library(ggplot2)
p<-ggplot(d_vis_3, aes(x=Var1,y=Freq,fill=Var2)) +
 geom_bar(position="dodge",stat='identity') + coord_flip()

print(p)

```

Employment Survey and Termination

```{r echo=FALSE, message=FALSE, warning=FALSE}


left_data<-subset(data_set,Termd==1)
stay_data<-subset(data_set,Termd==0)
ggplot() + geom_density(aes(x=EngagementSurvey), colour="red", data=left_data) + 
  geom_density(aes(x=EngagementSurvey), colour="blue", data=stay_data)

```


Employment Satisfaction and Termination

```{r echo=FALSE, message=FALSE, warning=FALSE}
left_data<-subset(data_set,Termd==1)
stay_data<-subset(data_set,Termd==0)
ggplot() + geom_density(aes(x=EmpSatisfaction), colour="red", data=left_data) + 
  geom_density(aes(x=EmpSatisfaction), colour="blue", data=stay_data)

```


#ProjectCount VS AverageMonthlyHours [BOXPLOT]
#Looks like the average employees who stayed worked about 200hours/month. Those that had a turnover worked about 250hours/month and 150hours/month
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(ggplot2)
data_set$SpecialProjectsCount
p<-ggplot(data_set, aes(x = factor(SpecialProjectsCount), y = Absences, fill = factor(Termd))) +
  geom_boxplot() + scale_fill_manual(values = c("yellow", "orange"))
print(p)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(data_set, aes(EngagementSurvey,salary, color = Termd)) +
  geom_point(shape = 16, size = 5, show.legend = FALSE) +
  theme_minimal() +
  scale_color_gradient(low = "#0091ff", high = "#f0650e")

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
library(Boruta)
data_set$Termd<-as.factor(data_set$Termd)
data_set <- na.omit(data_set)
boruta.train <- Boruta(Termd~., data = data_set, doTrace = 2)

print(boruta.train)
plot(boruta.train, xlab = "", xaxt = "n")
lz<-lapply(1:ncol(boruta.train$ImpHistory),function(i)
boruta.train$ImpHistory[is.finite(boruta.train$ImpHistory[,i]),i])
names(lz) <- colnames(boruta.train$ImpHistory)
Labels <- sort(sapply(lz,median))
axis(side = 1,las=2,labels = names(Labels),
at = 1:ncol(boruta.train$ImpHistory), cex.axis = 0.7)
           
```



```{r echo=FALSE, message=FALSE, warning=FALSE}
#Creating training and test sets for the logistic regression
smp_size <- floor(0.75 * nrow(data_set))

## set the seed to make your partition reproductible
set.seed(123)
train_ind <- sample(seq_len(nrow(data_set)), size = smp_size)

train <- data_set[train_ind, ]
test <- data_set[-train_ind, ]

dim(test)
dim(train)
```


Model

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(gmodels)
library (Hmisc)
library (caTools)
library (ROCR)
logit_model<-glm(Termd~+ManagerID+exp_in_company ,data=train,binomial())

summary(logit_model)

test$logit_model<-predict(logit_model,test)
#head(test)

colAUC(test$logit_model,test$Termd, plotROC=TRUE)


#Now using that threshold created the predicted values for each record
test$prediction<-ifelse(test$logit_model>=-.95,1,0)

#Make use of the confusion matrix to calculate accuracy, precision and recall
#CrossTable(test$left, test$prediction,prop.chisq=FALSE,prop.r=FALSE,prop.t=FALSE)
conf_mat<-table(test$Termd,test$prediction)

#print(conf_mat)
#class(conf_mat)
accuracy<-(conf_mat[1,1]+conf_mat[2,2])/(conf_mat[1,1]+conf_mat[2,2]+conf_mat[1,2]+conf_mat[2,1])
recall<-(conf_mat[2,2])/(conf_mat[1,2]+conf_mat[2,2])
precision<-(conf_mat[2,2])/(conf_mat[2,2]+conf_mat[2,1])

print(c("Accuracy:",accuracy))
print(c("Precision:",precision))
print(c("Recall:",recall))

```


caret
```{r echo=FALSE, message=FALSE, warning=FALSE}

# load the library
library(caret)
# load the iris dataset

# define training control
train_control <- trainControl(method="cv", number=10)
train$left<-as.factor(train$Termd)
# fix the parameters of the algorithm
grid <- expand.grid()
# train the model
model <- train(Termd~., data=train, trControl=train_control, method="glm",family=binomial())
#model <- train(left~satisfaction_level+last_evaluation+number_project+exp_in_company+average_montly_hours, data=train, trControl=train_control, method="glm",family=binomial())
# summarize results
print(model)
```
